---
title: "Statistical Sensitivity Analysis Tutorial"
author: "Brice Ozenne"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

# Introduction

This tutorial demonstrates a sensitivity analysis framework for multiverse analyses, following the approach described in the manuscript "A sensitivity analysis of preprocessing pipelines: Toward a solution for multiverse analyses". The framework uses statistical tools for assessing the heterogeneity of multiple preprocessing pipelines and testing hypotheses across them.

# Setup

Load the necessary R packages:

```{r}
library(mvtnorm)
library(Matrix)
library(LMMstar)
library(data.table)
library(ggplot2)
```

Check LMMstar version (should be 1.1.0)
```{r}
packageVersion("LMMstar")
```

# Data Simulation

We will first simulate data for a multiverse scenario.

```{r}
## ** parametrisation
n.obs <- 500 ## number of subjects in each group
beta <- 0.5 ## group difference
noise.Sigma <- as.matrix(bdiag((diag(0.05,15,15) + 0.95)*2.5, matrix(0.25), matrix(5), matrix(7.5), matrix(10), matrix(15))) ## noise distribution
n.pipe <- NCOL(noise.Sigma)

## ** simulate
set.seed(1)
X <- factor(c(rep("Male",n.obs),rep("Female",n.obs)))
dtW.data <- data.table(id = 1:(2*n.obs),
                       X = X,
                       X.num = as.numeric(X)-1,
                       Y = rnorm(2*n.obs, mean = beta*(as.numeric(X)-1), sd = 1)) ## Y is not observed
noise.pipeline <- rmvnorm(n=2*n.obs, sigma = noise.Sigma)
dtW.data <- cbind(dtW.data, matrix(dtW.data$Y, ncol = n.pipe, nrow = 2*n.obs, byrow = FALSE) + noise.pipeline)
names(dtW.data) <- c("id","X","X.num","Y",paste0("Ypip.",1:n.pipe))

## move to the long format (observed data)
dtL.data <- melt(dtW.data, id.vars = c("id","X","X.num"),
                 measure.vars = paste0("Ypip.",1:n.pipe),
                 value.name = "Ypip", variable.name = "pipeline")
```

# Analysis - fit pipeline specific models & get pipeline specific estimates

We run multiple linear models, one per pipeline (argument by), and extract the test w.r.t. the gender effect (argument effects). The argument repetition enables to keep track of the structure of the data to be able to perform statistical inference accounting for within-subject correlation.

```{r}
e.mlmm <- mlmm(Ypip ~ X, data = dtL.data,
               repetition = ~1|id,
               by = "pipeline",
               effects = "XMale=0")

summary(e.mlmm) ## summary result across all pipeline
```

```{r}
summary(e.mlmm$model[[1]])  ## fitted model for the first pipeline
```

## ** [optional] evaluate each pipeline in term of s.e. and how they correlate
```{r}
e.mlmm_vcov <- vcov(e.mlmm) 
sqrt(diag(e.mlmm_vcov)) ## standard error w.r.t. each pipeline
```
```{r}
range(cov2cor(e.mlmm_vcov)[lower.tri(e.mlmm_vcov)]) ## correlation between pipeline estimates
```
```{r}
plot(e.mlmm, "heat") ## automatic graphical display
```

## ** global effect estimator
```{r}
confint(e.mlmm, method = "average", columns = c("estimate","se","df","lower","upper","p.value"))
confint(e.mlmm, method = "pool.fixse", columns = c("estimate","se","df","lower","upper","p.value"))
confint(e.mlmm, method = "pool.gls", columns = c("estimate","se","df","lower","upper","p.value"))
confint(e.mlmm, method = "pool.gls1", columns = c("estimate","se","df","lower","upper","p.value"))
```
A more concise syntax is:
```{r}
model.tables(e.mlmm, method = c("average","pool.fixse","pool.gls","pool.gls1"))
```

# Visualization

Forest plot
```{r}
plot(e.mlmm) ## without global effect estimators
```

```{r}
plot(e.mlmm, method = c("none","average","pool.se","pool.gls","pool.gls1"))
```

Tune forest plot
```{r}
shape.forest <- c(rep(19,20),c(15,17,19,8))
color.forest <- c(rep("black",20), palette.colors()[2:5])
plot(e.mlmm, method = c("none","average","pool.se","pool.gls","pool.gls1"), 
     add.args = list(size.estimate = 2.5, size.ci = 1, width.ci = 0.5, shape = shape.forest, color = color.forest))
```

## ** proportion estimator
```{r}
proportion(e.mlmm, method = "single-step", n.sample = 0)
```

# Conclusion

This tutorial provides an overview of using R for sensitivity analysis of multiverse pipelines. For further details, refer to the original paper and the corresponding R code. If anything is not clear, or you would like to see additional scenarios added to this tutorial, we encourage you to open an issue here on GitHub.
